---
title: "FFADistance2.0"
author: "Jesse Garcia"
date: "March 4, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Using the R package Distance perform distance sampling and calculate abundance. 
```{r cars}
Winter2017FFAData <- read.csv("../Cleaned FFA Data Jan And Feb/FFA2017March_7_2017.csv", header = TRUE)

#Creating Perpendicular Distances from trail using Trigonometry
#Absolute Value of Distance * sin theta
Winter2017FFAData$PerpendicularDistances<- abs(Winter2017FFAData$Obs.Anml.Dist * sin(Winter2017FFAData$Angle*(3.14159/180)))
library(Distance)
summary(Winter2017FFAData$Obs.Anml.Dist)
#Removing Values greater than 100
Winter2017FFAData <- Winter2017FFAData[(Winter2017FFAData$Obs.Anml.Dist <= 100),] 
summary(Winter2017FFAData$Obs.Anml.Dist)
```

## Data Setup

Getting the three tables Ready as specified and required by Distance's manual and protocol.

#Region Table
```{r pressure, echo=TRUE}
#Region Label is 1 Because Bouamir Research Station is 1
#Area is 25000000 m2

regions <- data.frame(Region.Label = 1, Area= 25000000)

```


#Sample Table
```{r}
#There's 10 different trails
Region.Label=c(rep(1,10))

#Names of each trail
Sample.Label <- c("Bouamir", "Nest", "Elephant", "CDC", "Blowdown", "Entrance Trail", "RDA", "SB SWAMP LOOP","Jello", "Mbassako")

#Effort refers to the length of each trail. Multiplied by 4 because we conducted 4 surveys. 
Effort = c(4*960,4*1840,4*2305,4*2410,4*1375,4*3200,4*550,4*800,4*2100,4*1760)
samples <- data.frame(Region.Label=Region.Label, Sample.Label = Sample.Label, Effort = Effort)
```



#Creating a function that automates Distance analysis for our data and tables.

```{r}
require(Distance)
#Removing Things not on Trails
Winter2017FFAData <- Winter2017FFAData[!Winter2017FFAData$Repetition == 0,]
Winter2017FFAData <- Winter2017FFAData[!Winter2017FFAData$Trail == "Fruitwalk",]
Winter2017FFAData$Repetition = 1
UsingDistance <- function(Winter2017FFAData,Species,DistanceLimit) {
#Creating Perpendicular Distances using Trigonometry
#Absolute Value of Distance * sin theta
Winter2017FFAData$PerpendicularDistances<- abs(Winter2017FFAData$Obs.Anml.Dist * sin(Winter2017FFAData$Angle*(3.14159/180)))
#Removing Values greater than 100
Winter2017FFAData <- Winter2017FFAData[(Winter2017FFAData$PerpendicularDistances <= DistanceLimit),] 
BCHDistances <- Winter2017FFAData$PerpendicularDistances[Winter2017FFAData$Species == Species]
BCHObject <- 1:length(BCHDistances)
Region.Label <- Winter2017FFAData$Repetition[Winter2017FFAData$Species == Species ]
Sample.Label <- Winter2017FFAData$Trail[Winter2017FFAData$Species == Species ]
#Population Size 
size <- Winter2017FFAData$Number[Winter2017FFAData$Species == Species ]
    data <- data.frame(distance      = BCHDistances,
                       object        = BCHObject,
                       Region.Label  = Region.Label,
                       Sample.Label  = Winter2017FFAData$Trail[Winter2017FFAData$Species == Species ],
                       size = size)
model <-    ds(data = data, sample.table = samples, region.table = regions, dht.group = FALSE)
}
```


# Modeling probability functions for 2017 Survey Data. 


```{r}
AGPModel <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "AGP", 100)
GCMModel <- UsingDistance(Winter2017FFAData, Species = "GCM", 100)
CRGModel <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "CRG", 100)
WNG <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "WNG", 100)
GFP <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "GFP", 100)
BCH <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "BCH", 100)
MM <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "MM", 100)
APH <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "APH", 100)
RBDH <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "RBDH", 100)
WCH <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "WCH", 100)
GBT <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "GBT", 100)
WTH <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "WTH", 100)
GT <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "GT", 100)
#BWC Doesnt converge
#BWC <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "BWC", 100)
#VT Not In our Data
#VT <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "VT", 100)
#Chimp Not in our data set
#CHIMP <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "CHIMP", 100)
#GT Not in our Data Set
#GT <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "GT", 100)
PPH <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "PPH", 100)
#GVT Doesnt work
#GVT <- UsingDistance(Winter2017FFAData = Winter2017FFAData, Species = "GVT", 100)
```
## 2017 Data Results. Commented out models did not work for the reasons specified above
```{r}
summary(AGPModel)
summary(GCMModel)
summary(CRGModel)
summary(WNG)
summary(GFP)
summary(BCH)
summary(MM)
summary(APH)
summary(RBDH)
summary(WCH)
summary(GBT)
summary(WTH)
summary(GT)
#summary(BWC)
#summary(VT)
#summary(CHIMP)
#summary(GT)
summary(PPH)
#summary(GVT)
```

## DISTANCE Sampling For 1999 Data
```{r}
require(Distance)
Winter1999FFAData<-read.csv("../Cleaned FFA Data Jan And Feb/FFA1999March_7_2017.csv")
UsingDistanceOld <- function(Winter1999FFAData,Species,DistanceLimit) {
  #Creating Perpendicular Distances using Trigonometry
  #Absolute Value of Distance * sin theta
  Winter1999FFAData$PerpendicularDistances<- Winter1999FFAData$Perp
  #Removing Values greater than 100
  Winter1999FFAData <- Winter1999FFAData[(Winter1999FFAData$PerpendicularDistances <= DistanceLimit),] 
  BCHDistances <- Winter1999FFAData$PerpendicularDistances[Winter1999FFAData$Species == Species]
  BCHObject <- 1:length(BCHDistances)
  Region.Label <- 1
  Sample.Label <- Winter1999FFAData$Trail[Winter1999FFAData$Species == Species ]
  #Population Size 
  size <- Winter1999FFAData$Number[Winter1999FFAData$Species == Species ]
  data <- data.frame(distance      = BCHDistances,
                     object        = BCHObject,
                     Region.Label  = Region.Label,
                     Sample.Label  = Winter1999FFAData$Trail[Winter1999FFAData$Species == Species ],
                     size = size)
  model <-    ds(data = data, sample.table = samples, region.table = regions, dht.group = FALSE)
}


## Data Setup

#Getting the three tables Ready

#Region Table

#Region Label is 1 Because Bouamir Research Station is 1
#Area is 25000 m2


regions <- data.frame(Region.Label = 1, Area= 25000000)


#Sample Table

Region.Label=c(rep(1,10))
Sample.Label <- c("BOU", "N6", "ELE", "CDC", "BD", "ET", "RDA", "SB","JL", "MBA")
Effort = c(4*960,4*1840,4*2305,4*2410,4*1375,4*3200,4*550,4*800,4*2100,4*1760)
samples <- data.frame(Region.Label=Region.Label, Sample.Label = Sample.Label, Effort = Effort)

AGPModel <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "AGP", 100)
GCMModel <- UsingDistanceOld(Winter1999FFAData, Species = "GCM", 100)
CRGModel <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "CRG", 100)
WNG <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "WNG", 100)
GFP <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "GFP", 100)
BCH <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "BCH", 100)
MM <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "MM", 100)
APH <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "APH", 100)
RBDH <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "RBDH", 100)
WCH <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "WCH", 100)
GBT <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "GBT", 100)
WTH <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "WTH", 100)
GT <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "GT", 100)
BWC <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "BWC", 100)
VT <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "VT", 100)
CHIMP <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "CHIMP", 100)
GT <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "GT", 100)
PPH <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "PPH", 100)
GVT <- UsingDistanceOld(Winter1999FFAData = Winter1999FFAData, Species = "GVT", 100)



```


##Old Results
```{r}
summary(AGPModel)
summary(GCMModel)
summary(CRGModel)
summary(WNG)
summary(GFP)
summary(BCH)
summary(MM)
summary(APH)
summary(RBDH)
summary(WCH)
summary(GBT)
summary(WTH)
summary(GT)
summary(BWC)
summary(VT)
summary(CHIMP)
summary(GT)
summary(PPH)
summary(GVT)

```

