---
title: "FFABiodiversityHistogramEdits2"
author: "Jesse Garcia"
date: "March 15, 2017"
output: html_document
---

##Attempting Toms Suggested Second Edits
#Removing long first name and changing to abbreviations
#Make sure the scales match from 0 to 300 with  expand_limits(y=c(0,300))
#Remove titles on all of them 
#Remove Y axis on birds
#Remove Legend From Birds
#Move Legend on Primates to Top Right
#Remove Y axis label text on birds
#Merging Graphs

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Creating the Two Data Sets and Cleaning Them Up

```{r warning=FALSE, echo=FALSE}


#import in Past dataset

Edited_Past_Data_1_2_ <- read.csv("../Cleaned FFA Data Jan And Feb/FFA1999March_7_2017.csv")
table(Edited_Past_Data_1_2_$Species)

#renaming datset

PastData <-Edited_Past_Data_1_2_

#Removing Turacos
PastData <- PastData[!(PastData$Species == "GT"),]

#grouping data into primates and birds
PastData$PrimateBird <- 0
PastData$PrimateBird[(PastData$Species == "GVT" | PastData$Species == "VT" | PastData$Species == "AGP" | PastData$Species == "APH" | PastData$Species == "BCH" | PastData$Species == "GBT" | PastData$Species == "GFP" | PastData$Species == "GT" | PastData$Species == "PPH" | PastData$Species == "RBDH" | PastData$Species == "WCH" | PastData$Species == "WTH")]="Bird"
PastData$PrimateBird[PastData$PrimateBird == 0] <- "Primate"


#Giving Each observation its Scientific Name for Graph labels
#Asteriks are to Show statistical significance
PastData$ScientificName <- "Unknown"
PastData$ScientificName[PastData$Species == "AGP"] = "P. erithacus"
PastData$ScientificName[PastData$Species == "APH"] = "T. fasciatus" 
PastData$ScientificName[PastData$Species == "BCH"] = "C. atrata"
PastData$ScientificName[PastData$Species == "GBT"] = "C. cristata"
PastData$ScientificName[PastData$Species == "PPH"] = "B. fistulator"
PastData$ScientificName[PastData$Species == "RBDH"] = "T. camurus"
PastData$ScientificName[PastData$Species == "WCH"] = "T. albocristatus"
PastData$ScientificName[PastData$Species == "WTH"] = "B. albotibialis"
PastData$ScientificName[PastData$Species == "GFP"] = "T. calvus"

#Changing the Names of the  Primate species so they no longer are abbreviations. The asteriks Represent 
#Significant Differences

PastData$ScientificName[PastData$Species == "GCM"] = "L. albigena"
PastData$ScientificName[PastData$Species == "CRG"] = "C. pogonias" 
PastData$ScientificName[PastData$Species == "WNG"] = "C. nictitans"
PastData$ScientificName[PastData$Species == "MM"] = "C. cephus"
PastData$ScientificName[PastData$Species == "CHIMP"] = "P. troglodytes"
PastData$ScientificName[PastData$Species == "BWC"] = "C. guerza"

#Importing In Present Data set 
Winter2017FFAData <- read.csv("../Cleaned FFA Data Jan And Feb/FFA2017March_7_2017.csv")
#Data Has been Cleaned Using the Fixes Documented in The "Checking For Errors Document"
#Additionaly Two CRG observations were misrecorded as "GRG"" observations. 
#Fix: 
Winter2017FFAData$Species[(Winter2017FFAData$Species == "GRG")] = "CRG"
#Fix Number 2. "AGP" was recorded as "AGP " one time. Needs to be "AGP":
Winter2017FFAData$Species[(Winter2017FFAData$Species == "AGP ")] = "AGP"
#Table of how many times we recorded each of the different species. Not their abundance. 
table(Winter2017FFAData$Species)
#How many different species we found. There's 16.
unique(Winter2017FFAData$Species)
#Removing Fruit walk
Winter2017FFAData = Winter2017FFAData[!(Winter2017FFAData$Trail=="FW"),]
#Removing Things farther than 100 from Trail
Winter2017FFAData = Winter2017FFAData[Winter2017FFAData$Perp <= 100,]




#Removing Turacos
Winter2017FFAData <- Winter2017FFAData[!(Winter2017FFAData$Species == "GT"),]

#grouping data into primates and birds
Winter2017FFAData$PrimateBird <- 0
Winter2017FFAData$PrimateBird[(Winter2017FFAData$Species == "GVT" | Winter2017FFAData$Species == "VT" | Winter2017FFAData$Species == "AGP" | Winter2017FFAData$Species == "APH" | Winter2017FFAData$Species == "BCH" | Winter2017FFAData$Species == "GBT" | Winter2017FFAData$Species == "GFP" | Winter2017FFAData$Species == "GT" | Winter2017FFAData$Species == "PPH" | Winter2017FFAData$Species == "RBDH" | Winter2017FFAData$Species == "WCH" | Winter2017FFAData$Species == "WTH")]="Bird"
Winter2017FFAData$PrimateBird[Winter2017FFAData$PrimateBird == 0] <- "Primate"


#Giving Each observation its Scientific Name for Graph labels
#Asteriks are to Show statistical significance
Winter2017FFAData$ScientificName <- "Unknown"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "AGP"] = "P. erithacus"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "APH"] = "T. fasciatus" 
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "BCH"] = "C. atrata"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "GBT"] = "C. cristata"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "PPH"] = "B. fistulator"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "RBDH"] = "T. camurus"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "WCH"] = "T. albocristatus"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "WTH"] = "B. albotibialis"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "GFP"] = "T. calvus"

#Changing the Names of the  Primate species so they no longer are abbreviations. The asteriks Represent 
#Significant Differences

Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "GCM"] = "L. albigena"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "CRG"] = "C. pogonias" 
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "WNG"] = "C. nictitans"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "MM"] = "C. cephus"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "CHIMP"] = "P. troglodytes"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "BWC"] = "C. guerza"


#Labeling Old Data as Old, New Data as new

Winter2017FFAData$Year <- "2017"
PastData$Year <- "1999"

#Removing Columns that Can't be merged because they are of different units in the past and Present
ColumnsToRemovePresent <- names(Winter2017FFAData) %in% c("Time", "Distance.frm.Start")

ColumnsToRemovePast <- names(PastData) %in% c("Time", "Distance.frm.Start")

Winter2017FFAData <- Winter2017FFAData[,!ColumnsToRemovePresent]
PastData <- PastData[,!ColumnsToRemovePast]
```

#Merging Past And Present Data Sets

```{r}
require(dplyr)

MergedFFAData <- bind_rows(Winter2017FFAData, PastData)

table(MergedFFAData$Species)
table(MergedFFAData$Trail)
summary(MergedFFAData$Perp)

```

#Splitting Past and Present Data Sets Into Birds And Primates
```{r echo=FALSE}
#Seperating into Primate Data and Bird Data
MergedFFADataBirds <- MergedFFAData[MergedFFAData$PrimateBird == "Bird",]
MergedFFADataPrimates <- MergedFFAData[MergedFFAData$PrimateBird == "Primate",]
```





#Creating A Histogram Of Estimated Individual of Encounterd Primates Over Time

```{r echo=FALSE}
#Summing Number of Primate Species By  Year
NumberOfPrimatesByEncounter <- aggregate(Number ~ ScientificName + Year, MergedFFADataPrimates, sum)
#ATTEMPTING TO ORDER HISTOGRAM BY LOWEST TO HIGHEST ABUNDANCES OF PAST DATA. 
#First Get only Past Data
NumberOfPrimatesByEncounterPast <- NumberOfPrimatesByEncounter[NumberOfPrimatesByEncounter$Year == "1999",]
#Now set the levels of the Dataset we want to use, to be the Names of the ordered Past data from lowest number to highest. 
NumberOfPrimatesByEncounter$ScientificName <- factor(NumberOfPrimatesByEncounter$ScientificName, levels = NumberOfPrimatesByEncounterPast$ScientificName[order(NumberOfPrimatesByEncounterPast$Number)])  # notice the changed order of factor levels
#Graph
PrimateHistogram <- ggplot(NumberOfPrimatesByEncounter, aes(x = ScientificName, y = Number, fill = Year)) + geom_bar(stat = "identity", position  = "dodge") + ggtitle("") + ylab("Total Number of Individuals Encountered") + labs(fill = "Year", x = "Primate Species") +  theme(plot.title = element_text(color="black", size=14, face="bold.italic", hjust = 0.5)) + theme_bw() + theme(plot.title = element_text(hjust = 0.5),axis.text.x = italic.16.text)+ expand_limits(y=c(0,300)) + theme(legend.position="none")
ggsave("EncounterAbundancePrimates.png", width = 13, height = 8)

```


#Creating A Histogram of Estimated Individual of Encounterd Birds Over Time
```{r echo=FALSE}
#Summing Number of Bird Species By  Year
NumberOfBirdsByEncounters <- aggregate(Number ~ ScientificName + Year, MergedFFADataBirds, sum)

#ATTEMPTING TO ORDER HISTOGRAM BY LOWEST TO HIGHEST ABUNDANCES OF PAST DATA. Lol did it
#First Get only Past Data
NumberOfBirdsByEncountersPast <- NumberOfBirdsByEncounters[NumberOfBirdsByEncounters$Year == "1999",]
#Now set the levels of the Dataset we want to use, to be the Names of the ordered Past data from lowest number to highest. 
NumberOfBirdsByEncounters$ScientificName <- factor(NumberOfBirdsByEncounters$ScientificName, levels = NumberOfBirdsByEncountersPast$ScientificName[order(NumberOfBirdsByEncountersPast$Number)])  # notice the changed order of factor levels
#Graph
BirdHistogram <- ggplot(NumberOfBirdsByEncounters, aes(x = ScientificName, y = Number, fill = Year)) + geom_bar(stat = "identity", position  = "dodge") + ggtitle("") + ylab("") + labs(fill = "Year", x = "Bird Species") +  theme(plot.title = element_text(color="black", size=14, face="bold.italic", hjust = 0.5)) + theme_bw() + theme(plot.title = element_text(hjust = 0.5),axis.text.x = italic.16.text, axis.text.y = element_blank(), legend.position=c(.9,.8)) + expand_limits(y=c(0,300)) 
ggsave("EncounterAbundanceBirds.png", width = 13, height = 8)
```

#Merging Graphs
```{r}
require(cowplot)
plot1 <- PrimateHistogram
plot2 <- BirdHistogram
plot_grid(plot1, plot2, align='h', labels=c('A)', 'B)'))

```

