---
title: "Winter 1999 and 2017 FFA Analysis"
author: "Jesse Garcia"
date: "3/2/2017"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Creating the Two Data Sets and Cleaning Them Up

```{r warning=FALSE, echo=FALSE}


#import in Past dataset

Edited_Past_Data_1_2_ <- read.csv("../Cleaned FFA Data Jan And Feb/FFA1999March_7_2017.csv")
table(Edited_Past_Data_1_2_$Species)

#renaming datset

PastData <-Edited_Past_Data_1_2_

#Removing Turacos
PastData <- PastData[!(PastData$Species == "GT"),]

#grouping data into primates and birds
PastData$PrimateBird <- 0
PastData$PrimateBird[(PastData$Species == "GVT" | PastData$Species == "VT" | PastData$Species == "AGP" | PastData$Species == "APH" | PastData$Species == "BCH" | PastData$Species == "GBT" | PastData$Species == "GFP" | PastData$Species == "GT" | PastData$Species == "PPH" | PastData$Species == "RBDH" | PastData$Species == "WCH" | PastData$Species == "WTH")]="Bird"
PastData$PrimateBird[PastData$PrimateBird == 0] <- "Primate"


#Giving Each observation its Scientific Name for Graph labels
#Asteriks are to Show statistical significance
PastData$ScientificName <- "Unknown"
PastData$ScientificName[PastData$Species == "AGP"] = "Psittacus erithacus"
PastData$ScientificName[PastData$Species == "APH"] = "Tockus fasciatus" 
PastData$ScientificName[PastData$Species == "BCH"] = "Ceratogymna atrata"
PastData$ScientificName[PastData$Species == "GBT"] = "Corythaeola cristata"
PastData$ScientificName[PastData$Species == "PPH"] = "Bycanistes fistulator"
PastData$ScientificName[PastData$Species == "RBDH"] = "Tockus camurus"
PastData$ScientificName[PastData$Species == "WCH"] = "Tropicranus albocristatus"
PastData$ScientificName[PastData$Species == "WTH"] = "Bycanistes albotibialis"
PastData$ScientificName[PastData$Species == "GFP"] = "Treron calvus"

#Changing the Names of the  Primate species so they no longer are abbreviations. The asteriks Represent 
#Significant Differences

PastData$ScientificName[PastData$Species == "GCM"] = "Lophocebus albigena"
PastData$ScientificName[PastData$Species == "CRG"] = "Cercopithecus pogonias" 
PastData$ScientificName[PastData$Species == "WNG"] = "Cercopithecus nictitans"
PastData$ScientificName[PastData$Species == "MM"] = "Cercopithecus cephus"
PastData$ScientificName[PastData$Species == "CHIMP"] = "Pan troglodytes"
PastData$ScientificName[PastData$Species == "BWC"] = "Colobus guerza"

#Importing In Present Data set 
Winter2017FFAData <- read.csv("../Cleaned FFA Data Jan And Feb/FFA2017March_7_2017.csv")
#Data Has been Cleaned Using the Fixes Documented in The "Checking For Errors Document"
#Additionaly Two CRG observations were misrecorded as "GRG"" observations. 
#Fix: 
Winter2017FFAData$Species[(Winter2017FFAData$Species == "GRG")] = "CRG"
#Fix Number 2. "AGP" was recorded as "AGP " one time. Needs to be "AGP":
Winter2017FFAData$Species[(Winter2017FFAData$Species == "AGP ")] = "AGP"
#Table of how many times we recorded each of the different species. Not their abundance. 
table(Winter2017FFAData$Species)
#How many different species we found. There's 16.
unique(Winter2017FFAData$Species)
#Removing Fruit walk
Winter2017FFAData = Winter2017FFAData[!(Winter2017FFAData$Trail=="FW"),]
#Removing Things farther than 100 from Trail
Winter2017FFAData = Winter2017FFAData[Winter2017FFAData$Perp <= 100,]




#Removing Turacos
Winter2017FFAData <- Winter2017FFAData[!(Winter2017FFAData$Species == "GT"),]

#grouping data into primates and birds
Winter2017FFAData$PrimateBird <- 0
Winter2017FFAData$PrimateBird[(Winter2017FFAData$Species == "GVT" | Winter2017FFAData$Species == "VT" | Winter2017FFAData$Species == "AGP" | Winter2017FFAData$Species == "APH" | Winter2017FFAData$Species == "BCH" | Winter2017FFAData$Species == "GBT" | Winter2017FFAData$Species == "GFP" | Winter2017FFAData$Species == "GT" | Winter2017FFAData$Species == "PPH" | Winter2017FFAData$Species == "RBDH" | Winter2017FFAData$Species == "WCH" | Winter2017FFAData$Species == "WTH")]="Bird"
Winter2017FFAData$PrimateBird[Winter2017FFAData$PrimateBird == 0] <- "Primate"


#Giving Each observation its Scientific Name for Graph labels
#Asteriks are to Show statistical significance
Winter2017FFAData$ScientificName <- "Unknown"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "AGP"] = "Psittacus erithacus"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "APH"] = "Tockus fasciatus" 
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "BCH"] = "Ceratogymna atrata"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "GBT"] = "Corythaeola cristata"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "PPH"] = "Bycanistes fistulator"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "RBDH"] = "Tockus camurus"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "WCH"] = "Tropicranus albocristatus"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "WTH"] = "Bycanistes albotibialis"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "GFP"] = "Treron calvus"

#Changing the Names of the  Primate species so they no longer are abbreviations. The asteriks Represent 
#Significant Differences

Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "GCM"] = "Lophocebus albigena"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "CRG"] = "Cercopithecus pogonias" 
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "WNG"] = "Cercopithecus nictitans"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "MM"] = "Cercopithecus cephus"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "CHIMP"] = "Pan troglodytes"
Winter2017FFAData$ScientificName[Winter2017FFAData$Species == "BWC"] = "Colobus guerza"


#Labeling Old Data as Old, New Data as new

Winter2017FFAData$Year <- "2017"
PastData$Year <- "1999"

#Removing Columns that Can't be merged because they are of different units in the past and Present
ColumnsToRemovePresent <- names(Winter2017FFAData) %in% c("Time", "Distance.frm.Start")

ColumnsToRemovePast <- names(PastData) %in% c("Time", "Distance.frm.Start")

Winter2017FFAData <- Winter2017FFAData[,!ColumnsToRemovePresent]
PastData <- PastData[,!ColumnsToRemovePast]
```

#Merging Past And Present Data Sets

```{r}
require(dplyr)

MergedFFAData <- bind_rows(Winter2017FFAData, PastData)

table(MergedFFAData$Species)
table(MergedFFAData$Trail)
summary(MergedFFAData$Perp)

```

#Splitting Past and Present Data Sets Into Birds And Primates
```{r echo=FALSE}
#Seperating into Primate Data and Bird Data
MergedFFADataBirds <- MergedFFAData[MergedFFAData$PrimateBird == "Bird",]
MergedFFADataPrimates <- MergedFFAData[MergedFFAData$PrimateBird == "Primate",]
```


#Creating A Histogram Of Bird Encounters Over Time
```{r echo=FALSE}
#Creating a histogram of encounters
library(ggplot2)

##Attempting to Create Histogram Ordered by least Encountered Species in the past to most encountered species in the past

#Seperating Past And Present Data
MergedFFADataBirdsPast <- MergedFFADataBirds[MergedFFADataBirds$Year == "1999",]
#Summing Encounter Events by Species. Creating a "named vector"
SummarizedEncounterFFADataBirdsPast <- table(MergedFFADataBirdsPast$ScientificName)
#Finding the order of this named vector from smallest to largest
OrderOfPastEncounters <- order(SummarizedEncounterFFADataBirdsPast)
#Getting Names of Named vector ordered by encounter events!
OrderedBirdsPastEncounters <- names(SummarizedEncounterFFADataBirdsPast)[OrderOfPastEncounters]

#Factoring Merged Bird Data with new "levels" that are defined by our Order
MergedFFADataBirds$ScientificName <- factor(MergedFFADataBirds$ScientificName, levels = OrderedBirdsPastEncounters)

#Making a style for our graph to italicize Species Names
italic.16.text <- element_text(face = "italic", color = "black", size = 8)
## axis.text.x for x axis only

ggplot(MergedFFADataBirds, aes(fill=Year, x=ScientificName)) + geom_bar(position = "dodge") + ggtitle("Changes in Groups of Bird Encounter Events Over Time") + ylab("Number Of Groups of Bird Encounters ") + labs(fill = "Year", x = "Species") + theme(plot.title = element_text(color="black", size=14, face="bold.italic", hjust = 0.5)) + theme_bw() + theme(plot.title = element_text(hjust = 0.5),axis.text.x = italic.16.text)
```


#Creating a Histogram of Primate Encounters Over Time
```{r echo= FALSE}
#Creating a histogram of encounters
library(ggplot2)

#Seperating Primate Encounter into a past data frame so that histogram is order from lowest to highest encounter events of past
MergedFFADataPrimatesPast <- MergedFFADataPrimates[MergedFFADataPrimates$Year == "1999",]
#Getting Encounter events
SummarizedPastPrimateEncounters <- table(MergedFFADataPrimatesPast$ScientificName)
#Ordering Past Encounters by how often they show up in data set (lowest to highest)
OrderOfPastEncounters <- order(SummarizedPastPrimateEncounters)

#ATTEMPTING TO ORDER HISTOGRAM BY LOWEST TO HIGHEST ABUNDANCES OF PAST DATA. Lol did it
#First Get only Past Data
#Now set the levels of the Dataset we want to use, to be the Names of the ordered Past data from lowest number to highest. 
MergedFFADataPrimates$ScientificName <- factor(MergedFFADataPrimates$ScientificName, levels = names(SummarizedPastPrimateEncounters)[OrderOfPastEncounters])  # notice the changed order of factor levels

#Making a style for our graph to italicize Species Names
italic.16.text <- element_text(face = "italic", color = "black", size = 8)
## axis.text.x for x axis only




ggplot(MergedFFADataPrimates, aes(fill=Year, x=ScientificName)) + geom_bar(position = "dodge") + ggtitle("Changes in Groups of Primate Encounter Events Over Time") + ylab("Number Of Groups of Primate Encounters ") + labs(fill = "Year", x = "Species") + theme(plot.title = element_text(color="black", size=14, face="bold.italic", hjust = 0.5)) + theme_bw() + theme(plot.title = element_text(hjust = 0.5),axis.text.x = italic.16.text)
```

#Creating A Histogram Of Estimated Individual of Encounterd Primates Over Time

```{r echo=FALSE}
#Summing Number of Primate Species By  Year
NumberOfPrimatesByEncounter <- aggregate(Number ~ ScientificName + Year, MergedFFADataPrimates, sum)
#ATTEMPTING TO ORDER HISTOGRAM BY LOWEST TO HIGHEST ABUNDANCES OF PAST DATA. Lol did it
#First Get only Past Data
NumberOfPrimatesByEncounterPast <- NumberOfPrimatesByEncounter[NumberOfPrimatesByEncounter$Year == "1999",]
#Now set the levels of the Dataset we want to use, to be the Names of the ordered Past data from lowest number to highest. 
NumberOfPrimatesByEncounter$ScientificName <- factor(NumberOfPrimatesByEncounter$ScientificName, levels = NumberOfPrimatesByEncounterPast$ScientificName[order(NumberOfPrimatesByEncounterPast$Number)])  # notice the changed order of factor levels
#Graph
ggplot(NumberOfPrimatesByEncounter, aes(x = ScientificName, y = Number, fill = Year)) + geom_bar(stat = "identity", position  = "dodge") + ggtitle("Encounter Abundance of Primates 1999 vs 2017") + ylab("Total Number of Individuals Encountered") + labs(fill = "Year", x = "Species") +  theme(plot.title = element_text(color="black", size=14, face="bold.italic", hjust = 0.5)) + theme_bw() + theme(plot.title = element_text(hjust = 0.5),axis.text.x = italic.16.text)
ggsave("EncounterAbundancePrimates.png", width = 13, height = 8)

```


#Creating A Histogram of Estimated Individual of Encounterd Birds Over Time
```{r echo=FALSE}
#Summing Number of Bird Species By  Year
NumberOfBirdsByEncounters <- aggregate(Number ~ ScientificName + Year, MergedFFADataBirds, sum)

#ATTEMPTING TO ORDER HISTOGRAM BY LOWEST TO HIGHEST ABUNDANCES OF PAST DATA. Lol did it
#First Get only Past Data
NumberOfBirdsByEncountersPast <- NumberOfBirdsByEncounters[NumberOfBirdsByEncounters$Year == "1999",]
#Now set the levels of the Dataset we want to use, to be the Names of the ordered Past data from lowest number to highest. 
NumberOfBirdsByEncounters$ScientificName <- factor(NumberOfBirdsByEncounters$ScientificName, levels = NumberOfBirdsByEncountersPast$ScientificName[order(NumberOfBirdsByEncountersPast$Number)])  # notice the changed order of factor levels
#Graph
ggplot(NumberOfBirdsByEncounters, aes(x = ScientificName, y = Number, fill = Year)) + geom_bar(stat = "identity", position  = "dodge") + ggtitle("Encounter Abundance of Birds 1999 vs 2017") + ylab("Total Number of Individuals Encountered") + labs(fill = "Year", x = "Species") +  theme(plot.title = element_text(color="black", size=14, face="bold.italic", hjust = 0.5)) + theme_bw() + theme(plot.title = element_text(hjust = 0.5),axis.text.x = italic.16.text)
ggsave("EncounterAbundanceBirds.png", width = 13, height = 8)
```


#Exploratory Regression On Primates (EVERYTHING BElOW NOT INCLUDED IN PROJECT. ROUGH EXPLORATORY ANALYSIS)
```{r}

#Checking if Past Precent Predicts Current Percent
MergedFFADataPrimates <- aggregate(Number ~ ScientificName + Year, MergedFFADataPrimates, sum)
MergedFFADataPrimatesPast <- MergedFFADataPrimates[MergedFFADataPrimates$Year == "1999",]
MergedFFADataPrimatesPresent <- MergedFFADataPrimates[!MergedFFADataPrimates$Year == "1999",]
MergedFFADataPrimatesPastTotal <- sum(MergedFFADataPrimatesPast$Number)
MergedFFADataPrimatesPresentTotal <- sum(MergedFFADataPrimatesPresent$Number)
MergedFFADataPrimatesPast$PastPercent <- MergedFFADataPrimatesPast$Number/MergedFFADataPrimatesPastTotal
MergedFFADataPrimatesPresent$PresentPercent <-  MergedFFADataPrimatesPresent$Number/MergedFFADataPrimatesPresentTotal
require(dplyr)
MergedPercents <- bind_cols(MergedFFADataPrimatesPast , MergedFFADataPrimatesPresent)
PrimatePercents <- MergedPercents
plot(MergedPercents$PastPercent, MergedPercents$PresentPercent)
MergedPercentsModel <- lm(PastPercent~PresentPercent, MergedPercents)
summary(MergedPercentsModel)

#Plain number regression
MergedFFADataPrimates <- aggregate(Number ~ ScientificName + Year, MergedFFADataPrimates, sum)
MergedFFADataPrimatesPast <- MergedFFADataPrimates[MergedFFADataPrimates$Year == "1999",]
MergedFFADataPrimatesPresent <- MergedFFADataPrimates[!MergedFFADataPrimates$Year == "1999",]
MergedFFADataPrimatesPastTotal <- sum(MergedFFADataPrimatesPast$Number)
MergedFFADataPrimatesPresentTotal <- sum(MergedFFADataPrimatesPresent$Number)
MergedFFADataPrimatesPast$PastPercent <- MergedFFADataPrimatesPast$Number
MergedFFADataPrimatesPresent$PresentPercent <-  MergedFFADataPrimatesPresent$Number
require(dplyr)
MergedPercents <- bind_cols(MergedFFADataPrimatesPast , MergedFFADataPrimatesPresent)
plot(MergedPercents$PastPercent, MergedPercents$PresentPercent)
MergedPercentsModel <- lm(PastPercent~PresentPercent, MergedPercents)
summary(MergedPercentsModel)



```

#Regression With Birds
```{r}
#Checking if Past Precent Predicts Current Percent
MergedFFADataBirds <- aggregate(Number ~ ScientificName + Year, MergedFFADataBirds, sum)
MergedFFADataBirdsPast <- MergedFFADataBirds[MergedFFADataBirds$Year == "1999",]
MergedFFADataBirdsPresent <- MergedFFADataBirds[!MergedFFADataBirds$Year == "1999",]
MergedFFADataBirdsPastTotal <- sum(MergedFFADataBirdsPast$Number)
MergedFFADataBirdsPresentTotal <- sum(MergedFFADataBirdsPresent$Number)
MergedFFADataBirdsPast$PastPercent <- MergedFFADataBirdsPast$Number/MergedFFADataBirdsPastTotal
MergedFFADataBirdsPresent$PresentPercent <-  MergedFFADataBirdsPresent$Number/MergedFFADataBirdsPresentTotal
require(dplyr)
MergedPercents <- bind_cols(MergedFFADataBirdsPast , MergedFFADataBirdsPresent)
plot(MergedPercents$PastPercent, MergedPercents$PresentPercent)
MergedPercentsModel <- lm(PastPercent~PresentPercent, MergedPercents)
BirdPercents <- MergedPercents
summary(MergedPercentsModel)

#Plain number regression
MergedFFADataBirds <- aggregate(Number ~ ScientificName + Year, MergedFFADataBirds, sum)
MergedFFADataBirdsPast <- MergedFFADataBirds[MergedFFADataBirds$Year == "1999",]
MergedFFADataBirdsPresent <- MergedFFADataBirds[!MergedFFADataBirds$Year == "1999",]
MergedFFADataBirdsPastTotal <- sum(MergedFFADataBirdsPast$Number)
MergedFFADataBirdsPresentTotal <- sum(MergedFFADataBirdsPresent$Number)
MergedFFADataBirdsPast$PastPercent <- MergedFFADataBirdsPast$Number
MergedFFADataBirdsPresent$PresentPercent <-  MergedFFADataBirdsPresent$Number
require(dplyr)
MergedPercents <- bind_cols(MergedFFADataBirdsPast , MergedFFADataBirdsPresent)
plot(MergedPercents$PastPercent, MergedPercents$PresentPercent)
MergedPercentsModel <- lm(PastPercent~PresentPercent, MergedPercents)
summary(MergedPercentsModel)


```

#Merging Birds and Primates
```{r}

PrimatePercents
BirdPercents
MergedPercents <- MergedPercents[-c(15,4),]
plot(MergedPercents)
MergedPercentsModel <- lm(PresentPercent~PastPercent, MergedPercents)
summary(MergedPercents)
plot(MergedPercents)

MergedPercents
```
```{r}
ShannonIndexStats <- read.csv("ShannonIndexTtest.csv")

p <- ggplot(ShannonIndexStats, aes(fill=Year, y=Shannon_Index, x=Year))
limits <- aes(ymax = Shannon_Index + CI, ymin= Shannon_Index - CI)
p + geom_bar(position=dodge, stat = "identity") + geom_errorbar(limits, position=dodge, width=0.25) + labs(y="Shannon Index", x = "Time Period", title = "Shannon's Index Comparison (Encountered Abundances)") + theme(legend.title=element_blank()) 

par(mfrow=c(1,2))
ShannonIndexStats <- read.csv("ShannonIndexTtest2.csv")

p <- ggplot(ShannonIndexStats, aes(fill=Year, y=Shannon_Index, x=Year))
limits <- aes(ymax = Shannon_Index + CI, ymin= Shannon_Index - CI)
p + geom_bar(position=dodge, stat = "identity") + geom_errorbar(limits, position=dodge, width=0.25) + labs(y="Shannon Index", x = "Time Period", title = "Shannon's Index Comparison (Distance Estimated Abundances)") + theme(legend.title=element_blank()) 
```


